---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset 1 : Diwali Sales Data

```{r}
#| label: load-dataset
#| message: false
```

### Introduction to the Dataset

The Diwali Sales dataset offers a fascinating look at shopping behavior during Diwali, one of India’s most celebrated festivals. Shared by Saad Haroon on Kaggle, it includes sales transactions from a retail store during this vibrant period.

This dataset captures key details about customers, such as their age, gender, and location, along with information on what they bought and how much they spent. By exploring this data, we can uncover insights into consumer trends and preferences during the festive season, helping retailers better understand how to connect with their customers during this joyful time.

### Source of the Data

The dataset comes from Kaggle, a well-known platform where data enthusiasts and professionals share and explore various datasets. It was shared by Saad Haroon and focuses on sales data from a retail store during the vibrant Diwali festival in India.

### Collection Method

This data was collected during the bustling Diwali season, a time when shopping is at its peak in India. It likely includes records from the store’s transactions, capturing essential details about each purchase, such as who bought what and how much they spent.

While the dataset doesn’t specify the exact methods used for collection, we can imagine that it was gathered through the store’s point-of-sale (POS) systems. These systems automatically record sales along with customer details right at the checkout, providing a comprehensive view of consumer behavior during this festive time. This means that the dataset reflects real shopping patterns during a significant cultural celebration, making it rich with insights into how people celebrate Diwali through their purchases.

```{r}

library(dplyr)

data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-11-14/diwali_sales_data.csv')

glimpse(data)

```

### Description of the Observations

The Diwali sales dataset offers a fascinating look into consumer behavior during one of India’s most celebrated festivals. It contains valuable information about customers and their purchases at a retail store during the Diwali period. Here are some key insights:

Customer Demographics: The dataset includes details such as age, gender, marital status, and occupation. This demographic information helps us understand how different groups of people shop during this festive season.

Sales Metrics: It tracks the total amount each customer spent and the number of orders they placed, revealing trends in purchasing behavior. Given that Diwali often sees a surge in shopping, these metrics can highlight which groups are most engaged in the festivities.

Geographic Distribution: By including the state and geographic zone of each customer, the dataset allows us to explore regional differences in spending habits and preferences, adding depth to our understanding of cultural influences on shopping.

Product Categories: The dataset also categorizes the types of products purchased, helping to identify which items are popular during Diwali celebrations. This information can be particularly useful for retailers planning their inventory and marketing strategies.

### Ethical Concerns

When analyzing this dataset, it’s important to be mindful of several ethical considerations:

Data Privacy: The dataset includes identifiers like User_ID and Cust_name. We must ensure that this information is kept confidential to protect individuals' privacy. Anonymizing the data is crucial to prevent any potential misuse.

Informed Consent: It’s essential to think about whether customers were aware that their data would be used in this way. Understanding the context in which the data was collected can guide us in using it ethically.

Fair Representation: We need to be careful not to reinforce stereotypes or biases when analyzing demographic information like gender and age. Our findings should be presented fairly and thoughtfully, avoiding any narratives that could misrepresent these groups.

Impact on Stakeholders: Finally, we should consider how our analyses might affect those involved, including customers and the retail organization. The insights we gather should aim to improve business practices and customer experiences rather than exploit demographic trends.

### Research Question 1 : **How do customer demographics—specifically age group and gender—influence spending behavior during the Diwali sales period?**

#### Importance

Understanding how different age groups and genders spend during the Diwali season is crucial for retailers. This knowledge can help businesses tailor their marketing strategies and product offerings to better meet the needs and preferences of their customers. By identifying trends in spending behavior, retailers can create targeted promotions that resonate with specific demographics, ultimately driving sales and enhancing customer satisfaction during this festive time.

#### Topic Description

This research investigates how customer demographics—specifically age group and gender—affect spending behavior during the Diwali sales period. Understanding these relationships can help retailers tailor their marketing strategies and product offerings.

#### Hypotheses

1.  **H1:** Younger age groups will spend less than older age groups during the Diwali sales.

2.  **H2:** Female customers will make more orders than male customers, resulting in a higher total amount spent.

#### Types of Variables

-   **Amount spent:** Quantitative (e.g., in Indian Rupees)

-   **Age group:** Categorical (e.g., Youth, Adult, Senior)

-   **Gender:** Categorical (e.g., Male, Female)

-   **Total number of orders:** Quantitative (e.g., count of orders)

### Research Question 2 : **What role do geographic location and marital status play in shaping spending patterns during the Diwali sales period?**

#### Importance

Exploring the impact of geographic location and marital status on spending can provide valuable insights for retailers looking to connect with their customers more effectively. Different regions and demographic groups may have unique shopping behaviors and preferences, especially during a culturally significant time like Diwali. By understanding these dynamics, businesses can craft targeted marketing campaigns and tailor their product selections, ensuring they cater to the specific needs of various communities and enhance the overall shopping experience.

#### Research Topic Description

This research explores the influence of geographic location and marital status on spending patterns during the Diwali sales period. By understanding these dynamics, retailers can better target their promotions and products based on the demographic characteristics of different regions.

#### Hypotheses

1.  **H1:** Customers in urban zones will spend more on average than those in rural zones.

2.  **H2:** Married customers will spend more than single customers during the Diwali sales.

#### Types of Variables

-   **Geographic zone:** Categorical (e.g., Urban, Suburban, Rural)

-   **Marital status:** Categorical (e.g., Married, Single)

-   **Total amount spent:** Quantitative (e.g., in Indian Rupees)

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
